---
categories:
- ""
- ""
date: "2017-10-31T22:42:51-05:00"
description: Nullam et orci eu lorem consequat tincidunt vivamus et sagittis magna
  sed nunc rhoncus condimentum sem. In efficitur ligula tate urna. Maecenas massa
  sed magna lacinia magna pellentesque lorem ipsum dolor. Nullam et orci eu lorem
  consequat tincidunt. Vivamus et sagittis tempus.
draft: false
image: ""
keywords: ""
slug: tempus
title: Pre-Programme Assignment 
---

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
library(janitor)
```

# Task 2: `gapminder` country comparison

You have seen the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. To get a glimpse of the dataframe, namely to see the variable names, variable types, etc., we use the `glimpse` function. We also want to have a look at the first 20 rows of data.

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

Your task is to produce two graphs of how life expectancy has changed over the years for the `country` and the `continent` you come from.

I have created the `country_data` and `continent_data` with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "China") # just choosing Greece, as this is where I come from

continent_data <- gapminder %>% 
            filter(continent == "Asia")
```

First, create a plot of life expectancy over time for the single country you chose. Map `year` on the x-axis, and `lifeExp` on the y-axis. You should also use `geom_point()` to see the actual data points and `geom_smooth(se = FALSE)` to plot the underlying trendlines. You need to remove the comments **\#** from the lines below for your code to run.

```{r, lifeExp_one_country}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
    geom_point() +
    geom_smooth(se = FALSE)+
    NULL 

plot1
```

Next we need to add a title. Create a new plot, or extend plot1, using the `labs()` function to add an informative title to the plot.

```{r, lifeExp_one_country_with_label}
 plot1<- plot1 +
   labs(title = "Life expectancy over time in China  ",
       x = "Year ",
       y = "Life Expectancy ") +
   NULL


 plot1
```

Secondly, produce a plot for all countries in the *continent* you come from. (Hint: map the `country` variable to the colour aesthetic. You also want to map `country` to the `group` aesthetic, so all points for each country are grouped together).

```{r lifeExp_one_continent}
 ggplot(continent_data, mapping = aes(x = year  , y = lifeExp , colour= country  , group = country))+
   geom_point() + 
   geom_smooth(se = FALSE) +
   NULL
```

Finally, using the original `gapminder` data, produce a life expectancy over time graph, grouped (or faceted) by continent. We will remove all legends, adding the `theme(legend.position="none")` in the end of our ggplot.

```{r lifeExp_facet_by_continent}
 ggplot(data = gapminder , mapping = aes(x = year , y = lifeExp , colour= country))+
   geom_point() + 
   geom_smooth(se = FALSE) +
   facet_wrap(~continent) +
   theme(legend.position="none") + #remove all legends
   NULL
```

Given these trends, what can you say about life expectancy since 1952? Again, don't just say what's happening in the graph. Tell some sort of story and speculate about the differences in the patterns.

> Type your answer after this blockquote.

-   As for China, the life expectancy since 1952 has showed a rapid increase from below 45 to over 70. The growth rate is smaller after 1975.And I observed that the life expectancy decreased in 1962 compared to 1957, which is against the trend. This is probably result from the famine in the 1960s in China.
-   As for Asia,life expectancy varies greatly from country to country. Japan has the highest life expectancy which is over 80 now while Yeman has the lowest which is around 44 now. Most countries has risen rapidly since 1952 but Iraq showed a different trend because of war.
-   As for the difference between continent, Europe, Americas and Oceania show a milder growth compared with Asia and Africa, which indicates the rapid development in quality of life in developing countries and areas in the past 70 years. The other point is that there are many countries in Africa that have experienced retrogression in life expectancy, which indicates these countries faced serious economic challenges.

# Task 3: Brexit vote analysis

We will have a look at the results of the 2016 Brexit vote in the UK. First we read the data using `read_csv()` and have a quick glimpse at the data

```{r load_brexit_data, warning=FALSE, message=FALSE}
brexit_results <- read_csv(here::here("data","brexit_results.csv"))


glimpse(brexit_results)
```

The data comes from [Elliott Morris](https://www.thecrosstab.com/), who cleaned it and made it available through his [DataCamp class on analysing election and polling data in R](https://www.datacamp.com/courses/analyzing-election-and-polling-data-in-r).

Our main outcome variable (or y) is `leave_share`, which is the percent of votes cast in favour of Brexit, or leaving the EU. Each row is a UK [parliament constituency](https://en.wikipedia.org/wiki/United_Kingdom_Parliament_constituencies).

To get a sense of the spread, or distribution, of the data, we can plot a histogram, a density plot, and the empirical cumulative distribution function of the leave % in all constituencies.

```{r brexit_histogram, warning=FALSE, message=FALSE}

# histogram
h1 <-ggplot(brexit_results, aes(x = leave_share)) +
  geom_histogram(binwidth = 2.5)
h1+labs(title='The histogram of Percentage of votes that in favour of Brexit in all constituencies ',subtitle='histogram',x='The percentage that in favour of Brexit',y='The number of constituencies')

# density plot-- think smoothed histogram
d1<-ggplot(brexit_results, aes(x = leave_share)) +
  geom_density()
d1+labs(title='Distribution of Percentage of votes that in favour of Brexit in all constituencies ',subtitle='Density plot',x='the percentage that in favour of Brexit',y='density')


# The empirical cumulative distribution function (ECDF) 
c1<-ggplot(brexit_results, aes(x = leave_share)) +
  stat_ecdf(geom = "step", pad = FALSE) +
  scale_y_continuous(labels = scales::percent)
c1+labs(title='The empirical cumulative distribution of Percentage of votes that in favour of Brexit in all constituencies ',subtitle='ECDF',x='the percentage that in favour of Brexit',y='probability')


```

One common explanation for the Brexit outcome was fear of immigration and opposition to the EU's more open border policy. We can check the relationship (or correlation) between the proportion of native born residents (`born_in_uk`) in a constituency and its `leave_share`. To do this, let us get the correlation between the two variables

```{r brexit_immigration_correlation}
brexit_results %>% 
  select(leave_share, born_in_uk) %>% 
  cor()
```

The correlation is almost 0.5, which shows that the two variables are positively correlated.

We can also create a scatterplot between these two variables using `geom_point`. We also add the best fit line, using `geom_smooth(method = "lm")`.

```{r brexit_immigration_plot}
r1<-ggplot(brexit_results, aes(x = born_in_uk, y = leave_share)) +
  geom_point(alpha=0.3) +
  
  # add a smoothing line, and use method="lm" to get the best straight-line
  geom_smooth(method = "lm") + 
  
  # use a white background and frame the plot with a black box
  theme_bw() +
  NULL
r1+labs(title='The scatterplot of the proportion of native born residents and the tendency to vote for Brexit ',subtitle = 'scatterplot',x='the proportion of native born residents',y='the percetage of vote for Brexit')
```

You have the code for the plots, I would like you to revisit all of them and use the `labs()` function to add an informative title, subtitle, and axes titles to all plots.

What can you say about the relationship shown above? Again, don't just say what's happening in the graph. Tell some sort of story and speculate about the differences in the patterns.

> Type your answer after, and outside, this blockquote.

According to the scatterplot and the correlation which is 0.49, we can conclude that the proportion of native born residents in a constituency and its percentage of vote in favour of Brexit is *positively related*. In other words, the fear of immigration has *some kind of influence on* the result of Brexit. **But correlation is not equal to causality**. Thus, we can't come to conclusion that Brexit outcome result from fear of immigration and opposition to the EU's more open border policy. We still need more proof to examine this argument.

# Task 4: Animal rescue incidents attended by the London Fire Brigade

[The London Fire Brigade](https://data.london.gov.uk/dataset/animal-rescue-incidents-attended-by-lfb) attends a range of non-fire incidents (which we call 'special services'). These 'special services' include assistance to animals that may be trapped or in distress. The data is provided from January 2009 and is updated monthly. A range of information is supplied for each incident including some location information (postcode, borough, ward), as well as the data/time of the incidents. We do not routinely record data about animal deaths or injuries.

Please note that any cost included is a notional cost calculated based on the length of time rounded up to the nearest hour spent by Pump, Aerial and FRU appliances at the incident and charged at the current Brigade hourly rate.

```{r load_animal_rescue_data, warning=FALSE, message=FALSE}

url <- "https://data.london.gov.uk/download/animal-rescue-incidents-attended-by-lfb/8a7d91c2-9aec-4bde-937a-3998f4717cd8/Animal%20Rescue%20incidents%20attended%20by%20LFB%20from%20Jan%202009.csv"

animal_rescue <- read_csv(url,
                          locale = locale(encoding = "CP1252")) %>% 
  janitor::clean_names()


glimpse(animal_rescue)
```
One of the more useful things one can do with any data set is quick counts, namely to see how many observations fall within one category. For instance, if we wanted to count the number of incidents by year, we would either use `group_by()... summarise()` or, simply [`count()`](https://dplyr.tidyverse.org/reference/count.html)

```{r, instances_by_calendar_year}

animal_rescue %>% 
  dplyr::group_by(cal_year) %>% 
  summarise(count=n())

animal_rescue %>% 
  count(cal_year, name="count")

```

Let us try to see how many incidents we have by animal group. Again, we can do this either using group_by() and summarise(), or by using count()

```{r, animal_group_percentages}
animal_rescue %>% 
  group_by(animal_group_parent) %>% 
  
  #group_by and summarise will produce a new column with the count in each animal group
  summarise(count = n()) %>% 
  
  # mutate adds a new column; here we calculate the percentage
  mutate(percent = round(100*count/sum(count),2)) %>% 
  
  # arrange() sorts the data by percent. Since the default sorting is min to max and we would like to see it sorted
  # in descending order (max to min), we use arrange(desc()) 
  arrange(desc(percent))


animal_rescue %>% 
  
  #count does the same thing as group_by and summarise
  # name = "count" will call the column with the counts "count" ( exciting, I know)
  # and 'sort=TRUE' will sort them from max to min
  count(animal_group_parent, name="count", sort=TRUE) %>% 
  mutate(percent = round(100*count/sum(count),2))


```

Do you see anything strange in these tables? 

Answer:There are two animal groups both named cats.They should be comined together.

Finally, let us have a look at the notional cost for rescuing each of these animals. As the LFB says,

> Please note that any cost included is a notional cost calculated based on the length of time rounded up to the nearest hour spent by Pump, Aerial and FRU appliances at the incident and charged at the current Brigade hourly rate.

There is two things we will do:

1. Calculate the mean and median `incident_notional_cost` for each `animal_group_parent`
2. Plot a boxplot to get a feel for the distribution of `incident_notional_cost` by `animal_group_parent`.

Before we go on, however, we need to fix `incident_notional_cost` as it is stored as a `chr`, or character, rather than a number.

```{r, parse_incident_cost,message=FALSE, warning=FALSE}

# what type is variable incident_notional_cost from dataframe `animal_rescue`
typeof(animal_rescue$incident_notional_cost)

# readr::parse_number() will convert any numerical values stored as characters into numbers
animal_rescue <- animal_rescue %>% 

  # we use mutate() to use the parse_number() function and overwrite the same variable
  mutate(incident_notional_cost = parse_number(incident_notional_cost))

# incident_notional_cost from dataframe `animal_rescue` is now 'double' or numeric
typeof(animal_rescue$incident_notional_cost)

```

Now tht incident_notional_cost is numeric, let us quickly calculate summary statistics for each animal group. 


```{r, stats_on_incident_cost,message=FALSE, warning=FALSE}

animal_rescue %>% 
  
  # group by animal_group_parent
  group_by(animal_group_parent) %>% 
  
  # filter resulting data, so each group has at least 6 observations
  filter(n()>6) %>% 
  
  # summarise() will collapse all values into 3 values: the mean, median, and count  
  # we use na.rm=TRUE to make sure we remove any NAs, or cases where we do not have the incident cos
  summarise(mean_incident_cost = mean (incident_notional_cost, na.rm=TRUE),
            median_incident_cost = median (incident_notional_cost, na.rm=TRUE),
            sd_incident_cost = sd (incident_notional_cost, na.rm=TRUE),
            min_incident_cost = min (incident_notional_cost, na.rm=TRUE),
            max_incident_cost = max (incident_notional_cost, na.rm=TRUE),
            count = n()) %>% 
  
  # sort the resulting data in descending order. You choose whether to sort by count or mean cost.
  arrange(desc(mean_incident_cost))

```


Compare the mean and the median for each animal group. waht do you think this is telling us?
A:Except for squirrel,ferret and rabbit, median incident costs are much lower than the mean incident costs. This indicates that the distribution of the incident costs is generally positively skewed.

Anything else that stands out? Any outliers?

A:The minimum of incident cost of dogs is zero while it is 255 or 260 in other animals' group.So this is an outlier.

Finally, let us plot a few plots that show the distribution of incident_cost for each animal group.

```{r, plots_on_incident_cost_by_animal_group,message=FALSE, warning=FALSE}

# base_plot
base_plot <- animal_rescue %>% 
  group_by(animal_group_parent) %>% 
  filter(n()>6) %>% 
  ggplot(aes(x=incident_notional_cost))+
  facet_wrap(~animal_group_parent, scales = "free")+
  theme_bw()

base_plot + geom_histogram()
base_plot + geom_density()
base_plot + geom_boxplot()
base_plot + stat_ecdf(geom = "step", pad = FALSE) +
  scale_y_continuous(labels = scales::percent)



```

Which of these four graphs do you think best communicates the variability of the `incident_notional_cost` values? Also, can you please tell some sort of story (which animals are more expensive to rescue than others, the spread of values) and speculate about the differences in the patterns.

-   I think the boxplot best communicates the variability of the incident cost of different animal groups. 
-   The graph shows that horses are the most expensive to rescue, with a range between 250-1000. This indicates that the time rounded up to the nearest hour spent by Pump, Aerial and FRU appliances at the incident of horse are the longest generally. 
-   Based on median incident cost, the cheapest to rescue is heavy livestock animals. 
-   Cats are the most common animal to rescue with over three thousand incidents while cows are the least common animal to rescue with only eight incidents. 

# Submit the assignment

Knit the completed R Markdown file as an HTML document (use the "Knit" button at the top of the script editor window) and upload it to Canvas.

## Details

If you want to, please answer the following

-   Who did you collaborate with: n/a
-   Approximately how much time did you spend on this problem set: 3 HOURS
-   What, if anything, gave you the most trouble: Xcode on MAC
